TacticToe: Learning the Game of Tactics

### 2018-02-15 tactictoe_jv.tex
Tweak some of the terminology
e.g., proof script -> tactic proof; theory file -> script file.

I have not done this globally yet

The first line of this theory open modules (called structures in \sml). Each of
them contains a list of global identifiers which become directly accessible in
the rest of the theory.
A local identifier \texttt{LIST\_INDUCT\_TAC} is declared after, which is a
tactic that performs induction on lists. And even further, the theorem
\texttt{MAP\_APPEND} is proven.
The global tactic \texttt{STRIP\_TAC} first removes universal quantifiers. Then,
the goal is split into a base case and a inductive case. And finally, both of
theses cases are solved by \texttt{ASM\_REWRITE\_TAC [MAP, APPEND]}, which
rewrites assumptions with the help of theorems previously declared in this
theory.

We first parse the theory to extract proof scripts. Each of them is found in
the third argument of a call to \texttt{store\_thm}. The result of proof script
extraction for the running example in presented in Example~\ref{ex:r0}.

\begin{example}\label{ex:r0}(Proof script extraction)
\small
\begin{lstlisting}[language=SMLSmall]
STRIP_TAC THEN LIST_INDUCT_TAC THEN ASM_REWRITE_TAC [MAP, APPEND]
\end{lstlisting}
\end{example}

In the next phase, we globalize identifiers of the proof scripts.
Infix operators such as \texttt{THEN} need to be processed in a special way so
that they keep their infixity status after globalization. For simplicity,
the
globalization of infix operators is omitted in Example~\ref{ex:r1}.
In this example, the three main cases that can happen during the globalization
are depicted. The first one is the globalization of identifiers declared in
modules. The global identifiers \texttt{STRIP\_TAC} and
\texttt{ASM\_REWRITE\_TAC} are prefixed by their module \texttt{Tactic}. In
this way, they will still be interpretable whether \texttt{Tactic} was open or
not. The local identifier \texttt{LIST\_INDUCT\_TAC} is replaced by its
definition which happens to contain two global identifiers.
The previous paragraph describe the globalization for all identifiers except
local theorems.  We do not replace a local
theorem by its definition because we want to avoid unfolding proof scripts into
other proof scripts.
If the theorem is stored in the \holfour database available across \holfour
developments, we can obtain the theorem value by calling \texttt{DB.fetch}.
Otherwise the globalization process fails and the local theorem identifier is
kept unchanged. A recorded tactic with an unchanged local theorem as argument
is only interpretable inside the current theory.

@@

Running the globalized version of a proof script will have the exact same
effect as the original. But since we want to extract information from this
script in the form tactics and their input goals, we modify it further.
In particular, we need to define at which level we should record the tactics of
the proof scripts. The simplest idea would be to record all \sml subexpressions
of type tactics. However, it will damage the quality of our data by
dramatically increasing the number of tactic associated with a single goal.
Imagine a proof script of the form \texttt{A THEN B THEN C} then the tactics
\texttt{A}, \texttt{A THEN B} and \texttt{A THEN B THEN C} would be valid
advice for something close to their common input goal. The tactic
\texttt{A THEN B THEN C} is likely to be specific. In contrast, we can consider
the tactic \texttt{REPEAT A} that is repetitively calling the tactic \texttt{A}
until
\texttt{A} has no effect. In this situation, it is often preferable to record
only \texttt{REPEAT A} than calls to \texttt{A}. Otherwise during the
proof search, we would have to make a branches created after each call of
\texttt{A}.
To sum up, we would like to record only the most general tactics which
make the most progress on a goal. As a trade-off between these two objectives,
we split the proofs into tactic units.

\begin{definition}(Tactic unit)
A tactic unit is an \sml expression of type tactics that does not contain a
infix operator at its root.
\end{definition}

Because these tactic units are constructed from visual information present in
the proof script, they often represent what a human consider to be one step of
the proof.

@@

The function \texttt{save} writes the tactic-goal pair to disk increasing the
number of entries in our database of tactics. The function
\texttt{tactic\_of\_sml} interprets the \sml code \texttt{stac}. The tactic is
then applied to the input goal to replicate the original behavior.
After all modified theories are rebuild, each call to a wrapped tactic in a
proof script is recorded as a pair containing its globalized code and its
input goal.

>>>>

The first line of this script file opens modules (called structures in \sml). Each of
them contains a list of global identifiers which become directly accessible in
the rest of the script.
A local identifier \texttt{LIST\_INDUCT\_TAC} is declared next, which is a
tactic that performs induction on lists. Below that, the theorem
\texttt{MAP\_APPEND} is proven.
The global tactic \texttt{STRIP\_TAC} first removes universal quantifiers. Then,
the goal is split into a base case and a inductive case. Finally, both of
theses cases are solved by \texttt{ASM\_REWRITE\_TAC [MAP, APPEND]}, which
rewrites assumptions with the help of theorems previously declared in this
theory.

We first parse the script file to extract tactic proofs. Each of them is found in
the third argument of a call to \texttt{store\_thm}. The result of tactic proof
extraction for the running example in presented in Example~\ref{ex:r0}.

\begin{example}\label{ex:r0}(Tactic proof extraction)
\small
\begin{lstlisting}[language=SMLSmall]
STRIP_TAC THEN LIST_INDUCT_TAC THEN ASM_REWRITE_TAC [MAP, APPEND]
\end{lstlisting}
\end{example}

In the next phase, we globalize identifiers of the tactic proofs.
Infix operators such as \texttt{THEN} need to be processed in a special way so
that they keep their infixity status after globalization. For simplicity,
the
globalization of infix operators is omitted in Example~\ref{ex:r1}.
In this example, the three main cases that can happen during the globalization
are depicted. The first one is the globalization of identifiers declared in
modules. The global identifiers \texttt{STRIP\_TAC} and
\texttt{ASM\_REWRITE\_TAC} are prefixed by their module \texttt{Tactic}. In
this way, they will still be interpretable whether \texttt{Tactic} was open or
not. The local identifier \texttt{LIST\_INDUCT\_TAC} is replaced by its
definition which happens to contain two global identifiers.
The previous paragraph describe the globalization for all identifiers except
local theorems.  We do not replace a local
theorem by its definition because we want to avoid unfolding tactic proofs into
other tactic proofs.
If the theorem is stored in the \holfour database available across \holfour
developments, we can obtain the theorem value by calling \texttt{DB.fetch}.
Otherwise the globalization process fails and the local theorem identifier is
kept unchanged. A recorded tactic with an unchanged local theorem as argument
is only interpretable inside the current theory.

@@

Running the globalized version of a tactic proof will have the exact same
effect as the original. But since we want to extract information from this
script in the form tactics and their input goals, we modify it further.
In particular, we need to define at which level we should record the tactics in
the tactic proof. The simplest idea would be to record all \sml subexpressions
of type \texttt{tactic}. However, it will damage the quality of our data by
dramatically increasing the number of tactics associated with a single goal.
Imagine a proof script of the form \texttt{A THEN B THEN C}; then the tactics
\texttt{A}, \texttt{A THEN B} and \texttt{A THEN B THEN C} would be valid
advice for something close to their common input goal. The tactic
\texttt{A THEN B THEN C} is likely to be specific. In contrast, we can consider
the tactic \texttt{REPEAT A} that is repetitively calling the tactic \texttt{A}
until
\texttt{A} has no effect. In this situation, it is often preferable to record
only \texttt{REPEAT A} than calls to \texttt{A}. Otherwise during the
proof search, we would have to make a branches created after each call of
\texttt{A}.
To sum up, we would like to record only the most general tactics which
make the most progress on a goal. As a trade-off between these two objectives,
we split proofs into tactic units.

\begin{definition}(Tactic unit)
A tactic unit is an \sml expression of type \texttt{tactic} that does not contain a
infix operator at its root.
\end{definition}

Because these tactic units are constructed from visual information present in
the tactic proof, they often represent what a human consider to be one step of
the proof.

@@

The function \texttt{save} writes the tactic-goal pair to disk increasing the
number of entries in our database of tactics. The function
\texttt{tactic\_of\_sml} interprets the \sml code \texttt{stac}. The tactic is
then applied to the input goal to replicate the original behavior.
After all modified theories are rebuild, each call to a wrapped tactic in a
tactic proof is recorded as a pair containing its globalized code and its
input goal.
